Home Nighttime Cough Monitoring

### 2025-01-10 Lilah Melzer/cough_monitor_analysis.qmd
cough monitor zero inflated models

```{r setup}
#| include: false
library(Hmisc)
library(redcapAPI)
library(tidyverse)
library(rspiro)
library(childsds)
library(arsenal)
library(lmerTest)
library(performance)
```

@@

# Those with end date after the last data pull will just end on the date of the
# pull
pull_date <- mdy("8/28/24")

@@

  "/Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF/Lilah Melzer/Lilah Cough Monitor/Data_Raw/Cough_w_TotalCounts_0824 - Cough_agg_w_count_0828.csv",

@@

@@

write.csv(agg_counts, file = "/Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF/Lilah Melzer/Lilah Cough Monitor/Data_Clean/cleaned_aggregated_counts.csv", na = "", row.names = F)
# Per night
per_night <- agg_counts %>%
  summarise(
    median_coughs = median(Coughs, na.rm = T),
    nights_data = sum(!is.na(Coughs))
demo$sid <- as.numeric(demo$sid)
demo <- left_join(demo, per_night)
label(demo$median_coughs) <- "Median Coughs per Night"
label(demo$nights_data) <- "Nights With Cough Data"

@@

@@

# Data for ATS abstract

```{r results='asis'}
per_person <- agg_counts %>%
  group_by(sid, Diagnosis) %>%
  summarise(
    `Nights` = n(),
    `Nights With Data` = sum(!is.na(Coughs)),
    `Prop. Nights With Data` = sum(!is.na(Coughs)) / n(),
    `Median Coughs per Night` = median(Coughs, na.rm = T),
    .groups = "drop"
  )
t2 <- tableby(
  Diagnosis ~ `Nights` + `Nights With Data` + `Prop. Nights With Data` +
    `Median Coughs per Night`,
  data = per_person, control = mycontrols
)
summary(t2, pfootnote = T)
```

@@

```{r}
#| label: fig-coughs-by-night
#| fig-cap: Coughs by nights from start
#| warning: false
ggplot(agg_counts, aes(
  x = Night, y = Coughs, group = factor(sid), color = Diagnosis
)) +
  geom_point(alpha = 0.1) +
  geom_line(alpha = 0.1) +
  geom_smooth(aes(group = Diagnosis))
```

## Model

```{r}
mod <- glmer(Coughs ~ Night + Diagnosis + (Night | sid),
  data = agg_counts, family = "poisson"
)
gtsummary::tbl_regression(mod, exponentiate = T)
# check_model(mod, re_formula = NA, residual_type = "normal")
```

@@

```{r}
mod <- glmer(Coughs ~ Night + Diagnosis + Sick + (Night | sid),
  data = agg_counts, family = "poisson"
)
gtsummary::tbl_regression(mod, exponentiate = T)
# check_model(mod, re_formula = NA, residual_type = "normal")
```

>>>>

```{r setup}
#| include: false
library(Hmisc)
library(redcapAPI)
library(tidyverse)
library(rspiro)
library(childsds)
library(arsenal)
library(lmerTest)
library(performance)
library(GLMMadaptive)
```

@@

# Last date should be either the end date in REDCap or the last day with data
# I
pull_date <- mdy("11/05/24")

@@

  "/Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF/Lilah Melzer/Lilah Cough Monitor/Data_Raw/Cough_w_TotalCounts_1105_NMDcorrected - Cough_agg_w_count_122324.csv",

@@

agg_counts$`#nights-monitoring` <- NULL

@@

@@

# Exclude participant 3 per Lilah
agg_counts <- agg_counts %>% filter(sid != 3)
write.csv(agg_counts,
  file = "/Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF/Lilah Melzer/Lilah Cough Monitor/Data_Clean/cleaned_aggregated_counts.csv",
  na = "", row.names = F

@@

# Summary data

```{r results='asis'}
per_person <- agg_counts %>%
  group_by(sid, Diagnosis) %>%
  summarise(
    `Nights` = n(),
    `Nights With Data` = sum(!is.na(Coughs)),
    `Prop. Nights With Data` = sum(!is.na(Coughs)) / n(),
    `Median Coughs per Night` = median(Coughs, na.rm = T),
    .groups = "drop"
  )
# Exclude the people with 0 nights of data
per_person <- per_person %>% filter(`Nights With Data` != 0)
t2 <- tableby(
  Diagnosis ~ `Nights` + `Nights With Data` + `Prop. Nights With Data` +
    `Median Coughs per Night`,
  data = per_person, control = mycontrols
)
summary(t2, pfootnote = T)
```

@@

```{r}
#| label: fig-coughs-by-night
#| fig-cap: Coughs by nights from start
#| warning: false
ggplot(agg_counts[!is.na(agg_counts$Coughs), ], aes(
  x = Night, y = Coughs, group = factor(sid), color = Diagnosis
)) +
  geom_point(alpha = 0.1) +
  geom_line(alpha = 0.1) +
  geom_smooth(aes(group = Diagnosis))
```

## Zero-inflated models

```{r}
zip1 <- mixed_model(Coughs ~ Night + Diagnosis,
  random = ~ 1 | sid, data = agg_counts,
  family = zi.poisson(), zi_fixed = ~1,
  zi_random = ~ 1 | sid
)
zip2 <- update(zip1, zi_fixed = ~ Night + Diagnosis)

anova(zip1, zip2)

zinb1 <- update(zip2, family = zi.negative.binomial())

anova(zip2, zinb1)

summary(zinb1)
# check_model(zinb1, re_formula = NA, residual_type = "normal")
```

@@

```{r}
zinb2 <- update(zinb1,
  fixed = Coughs ~ Night + Diagnosis + Sick,
  zi_fixed = ~ Night + Diagnosis + Sick
)
summary(zinb2)
# check_model(zinb2, re_formula = NA, residual_type = "normal")
```
